More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-07-02 2016-Weak-FH-Bisim/WeakBisim.tex
updated date
Petits nettoyages Eric

%\TODO{Add somewhere a subsection ``running example'' ?}

@@

\begin{figure}[bh]
\begin{mathpar}
  \small
  \inferrule
    {\inferrule
        {0 \xrightarrow {\delta}_{C_3} 1}
        {C_3
          \models
            \openrule{
              0 \xrightarrow {\delta}_{C_3} 1,\,
              \{\xrightarrow{\delta(x_1)}_P\},\,
              v_1=\delta(x_1)}
                      {\ostate{0}\OTarrow{v_1}\ostate{1}}
        }\\
%      \sm{fresh}v         \\
      \inferrule%*[right={L_1}]
        {
%          \sm{fresh}{a_Q} \\
          \inferrule
              {0 \xrightarrow{l}_{C_4} 0}
              % {\ostate{0}\xrightarrow{l}\ostate{0}}
              {C_4
                \models
                \openrule{
                      0 \xrightarrow l_{C_4} 0,\,
                      \sm{Pred}_{C_4}}
                      {\ostate{0}\OTarrow{l}\ostate{0}}
              }
        }
        {
          \textrm{Q>>R}\models
              \openrule
                  { 0 \xrightarrow {l}_{C_4} 0,\,
                    \{\xrightarrow{acc(x_2)}_Q\},\,
                    \ v_2=acc(x_2)
                  }
                  {\ostate{0}\OTarrow{v_2}\ostate{0}}
        }   
    }
    {
     \textrm{P>>(Q>>R)}
     \models
     \openrule
         { 0 \xrightarrow{\delta}_{C_3} 1, \\ 0 \xrightarrow{l}_{C_4} 0,\\
           \{\xrightarrow{\delta(x)}_P,\,\xrightarrow{acc(x)}_Q\}, \\
            a_3=v_1 \wedge v=a_3 \wedge x_1=x_2
           }
         {\ostate{00} \OTarrow{v} \ostate{10}}
      }\vspace{-4ex}
\end{mathpar}
  \caption{Proof of transition $OT_2$ (with interaction of processes $P$ and $Q$) for 
  ``P>>(Q>>R)'' \TODO{remove?}}
  \label{usingrules:OT2}
\end{figure}

@@

\TODO{Eric: refine the algorithm description - \ERIC{Done, can you check ?}}

@@

\subsection{Post-processing: Elimination of intermediate variables}

While building the open transitions, many new variable names are
created, as fresh local variables of synchronisation vectors, and
fresh result variables of intermediate OTs. While these are required
for the semantic construction, they are not significant in the
resulting open automaton. In fact they carry information on the
internal structure of the pNet system, and this information should
\_not\_ be visible when looking at the system as a black box, and in
particular when comparing its behaviour with another system, e.g. by
bisimulation.

In the resulting predicate the only significant variables are :\\
- the input variables of pLTS transitions \TODO{heu, non, meme pas, si
  ce sont des comm internes!}\\
- the actions of holes\\
- the result action at toplevel.

So we transform the predicates of the resulting OTs, eliminating all
intermediate variables. \TODO{Explain why this should work !! Maybe
  need to be more precise on unification in the matching...}

@@

\RAB{We let $\gamma$ range over sets of words and use $\dotcup$ as the union that appends sets of words.}

@@

We define $\vis{\overline\beta}$ as the mapping $\overline\beta$  without  invisible actions but where each element is now a list (if length 1).\\ $\vis{\beta_i^{i\in I}} = [\beta_i]^{i\in I'}$ s.t. $I'=\left\{i| i\in I \land \beta_i\neq \tau\right\}$.
\begin{definition}[Weak open transition]\label{def:weakOT}
A weak open transition over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
\begin{mathpar}
 \openrule
         {
           \gamma_j^{j\in J},\symb{True},Id(s)}
         {s \OTWeakarrow {\alpha} s'}
 \end{mathpar}
	Where $s, s'\in\mathcal{S}$ and $\gamma_j$
        is a list of transitions of the hole $j$, with each element of the list in $\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action
        of this open transition. \Pred\ and \Post\ are defined similarly to Definition~\ref{def:OT}. We use $\WT$ to range over sets of weak open transitions.

@@

\TODO{ the formal view is not much interesting}

\begin{property}[Weak open transition transitivity]
Let $A$ be an open automaton, and $<J,\mathcal{S},s_0,\WT>$ a weak open automaton derived from $A$.
Suppose we have the following weak open transitions in $\WT$:
\[\openrule
         {
           \overline{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1}
\qquad 
\openrule
         {
           \overline{\gamma_2},\Pred_2,\Post_2   }
         {s_1 \OTWeakarrow {\alpha} s_2}
\qquad
\openrule
         {
           \overline{\gamma_3},\Pred_3,\Post_3   }
         {s_2 \OTWeakarrow {\tau} s_3}
\]
Then there is a weak  open transition of the following form in $\WT$:
\[\openrule
         {
           \overline{\gamma},\Pred,\Post   }
         {s \OTWeakarrow {\alpha} s_3}
\]
Suppose additionally $\Post_3=({x_k\gets e_k})^{k\in K}$, we have the following
\[
\begin{array}{rcl}
\gamma&=& \gamma_1\dotcup\gamma_2\dotcup\gamma_3\\
\Pred&=&\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}\\
\Post &=& ({x_k\gets ((e_k\subst{\Post_2})\subst{\Post_1})})^{k\in K}
\end{array}\]
\end{property}

@@

\subsection{Weak bisimulation for open pNets}
\begin{definition}[Non-observability of silent actions]~\\
A pNet $\mylangle \pNet_i^{i\in I} , \Sort_j^{j\in J}, \overline{\symb{SV}}\myrangle$
 \emph{cannot observe silent actions} if it verifies:\\ $\forall i\in I\uplus J.\, \SV{(i\mapsto \tau)}{\tau}{\symb{True}}\in \overline{\symb{SV}}$ and $\forall i,j\in I\uplus J.\,i\neq j \Rightarrow \alpha_i\neq\tau$
\end{definition}

@@

\begin{definition}[Notation: Semantics of pNets as a weak open automaton]
Let $A$ be the open automaton expressing the semantics of an open \pNet: $\pNet$; let $<J,\mathcal{S},s_0,  \WT>$ be the weak open automata derived from $A$. For any $\WOT\in\WT$, we denote $\pNet \models \WOT$, this defines the semantics of the pNet \pNet as a weak open automaton.
\end{definition}

\TODO{move text below}

\begin{property}[Decomposition of weak open transitions] \TODO{TODO}
Let $A$ be the open automaton that is the semantics of a \pNet $P$, and $<J,\mathcal{S},s_0,\WT>$ a weak open automaton derived from $A$.
Suppose we have the following weak open transitions in $\WT$:
\[\openrule
         {
           \overline{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1}
\qquad 
\openrule
         {
           \overline{\gamma_2},\Pred_2,\Post_2   }
         {s_1 \OTWeakarrow {\alpha} s_2}
\qquad
\openrule
         {
           \overline{\gamma_3},\Pred_3,\Post_3   }
         {s_2 \OTWeakarrow {\tau} s_3}
\]
Then there is ag weak  open transition of the following form in $\WT$:
\[\openrule
         {
           \overline{\gamma},\Pred,\Post   }
         {s \OTWeakarrow {\alpha} s_3}
\]
Suppose additionally $\Post_3=({x_k\gets e_k})^{k\in K}$, we have the following
\[
\begin{array}{rcl}
\gamma&=& \gamma_1\dotcup\gamma_2\dotcup\gamma_3\\
\Pred&=&\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}\\
\Post &=& ({x_k\gets ((e_k\subst{\Post_2})\subst{\Post_1})})^{k\in K}
\end{array}\]
\end{property}

\TODO{We will discuss this corollary and the title.}
\begin{corollary}[Weak Composability]
	Consider two weak FH-bisimilar pNets with an arbitrary number of holes, when replacing, 
	inside those two original pNets, a subset of the holes by FH-bisimilar pNets, we 
	obtain two weak FH-bisimilar pNets.
\end{corollary}

@@

\TODO{Here we should include:
  \\  - A sufficient condition for an open PNet to preserve Weak-FH,
  including: 1) no non-deterministic Tau-choice in the pLTSs, 2) a
  synch vector propagating Tau from a subnet $pNet_i$ mandatory
  everytime $pNet_i$ can do some visible action in the current
  context, 3) if we have a vector <Tau, -, a, - > --> b with ``a'' a visible
  action of some CLOSED subnet, then b=Tau. 
  \\  - the corresponding Congruence theorem
  \\- An example
}

>>>>

@@

\begin{figure}[bh]
\begin{mathpar}
  \small
  \inferrule
    {\inferrule
        {0 \xrightarrow {\delta}_{C_3} 1}
        {C_3
          \models
            \openrule{
              0 \xrightarrow {\delta}_{C_3} 1,\,
              \{\xrightarrow{\delta(x_1)}_P\},\,
              v_1=\delta(x_1)}
                      {\ostate{0}\OTarrow{v_1}\ostate{1}}
        }\\
%      \sm{fresh}v         \\
      \inferrule%*[right={L_1}]
        {
%          \sm{fresh}{a_Q} \\
          \inferrule
              {0 \xrightarrow{l}_{C_4} 0}
              % {\ostate{0}\xrightarrow{l}\ostate{0}}
              {C_4
                \models
                \openrule{
                      0 \xrightarrow l_{C_4} 0,\,
                      \sm{Pred}_{C_4}}
                      {\ostate{0}\OTarrow{l}\ostate{0}}
              }
        }
        {
          \textrm{Q>>R}\models
              \openrule
                  { 0 \xrightarrow {l}_{C_4} 0,\,
                    \{\xrightarrow{acc(x_2)}_Q\},\,
                    \ v_2=acc(x_2)
                  }
                  {\ostate{0}\OTarrow{v_2}\ostate{0}}
        }   
    }
    {
     \textrm{P>>(Q>>R)}
     \models
     \openrule
         { 0 \xrightarrow{\delta}_{C_3} 1, \\ 0 \xrightarrow{l}_{C_4} 0,\\
           \{\xrightarrow{\delta(x)}_P,\,\xrightarrow{acc(x)}_Q\}, \\
            a_3=v_1 \wedge v=a_3 \wedge x_1=x_2
           }
         {\ostate{00} \OTarrow{v} \ostate{10}}
      }\vspace{-4ex}
\end{mathpar}
  \caption{Proof of transition $OT_2$ (with interaction of processes $P$ and $Q$) for 
  ``P>>(Q>>R)'' \TODO{Change for the new example or remove?}}
  \label{usingrules:OT2}
\end{figure}

@@

% \TODO{Eric: refine the algorithm description - \ERIC{Done, can you check ?}}

@@

%% \subsection{Post-processing: Elimination of intermediate variables}

%% While building the open transitions, many new variable names are
%% created, as fresh local variables of synchronisation vectors, and
%% fresh result variables of intermediate OTs. While these are required
%% for the semantic construction, they are not significant in the
%% resulting open automaton. In fact they carry information on the
%% internal structure of the pNet system, and this information should
%% \_not\_ be visible when looking at the system as a black box, and in
%% particular when comparing its behaviour with another system, e.g. by
%% bisimulation.

%% In the resulting predicate the only significant variables are :\\
%% - the input variables of pLTS transitions \TODO{heu, non, meme pas, si
%%   ce sont des comm internes!}\\
%% - the actions of holes\\
%% - the result action at toplevel.

%% So we transform the predicates of the resulting OTs, eliminating all
%% intermediate variables. \TODO{Explain why this should work !! Maybe
%%   need to be more precise on unification in the matching...}

@@

\RAB{We let $\gamma$ range over words (of action terms) and use $\dotcup$ as the union that appends words.}

@@

We define $\vis{\overline\beta}$ as the mapping $\overline\beta$  without  invisible actions but where each element is now a list (if length 1).\\ $\vis{\beta_i^{i\in I}} = [\beta_i]^{i\in I'}$ s.t. $I'=\left\{i| i\in I \land \beta_i\neq \tau\right\}$.
\begin{definition}[Weak open transition]\label{def:weakOT}
A weak open transition over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
\begin{mathpar}
 \openrule
         {
           \gamma_j^{j\in J},\Pred,\Post}
         {s \OTWeakarrow {\alpha} s'}
 \end{mathpar}
	Where $s, s'\in\mathcal{S}$ and $\gamma_j$
        is a list of transitions of the hole $j$, with each element of the list in $\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action
        of this open transition. \Pred\ and \Post\ are defined similarly to Definition~\ref{def:OT}. We use $\WT$ to range over sets of weak open transitions.

@@

%% \TODO{ the formal view is not much interesting}

%% \begin{property}[Weak open transition transitivity]
%% Let $A$ be an open automaton, and $<J,\mathcal{S},s_0,\WT>$ a weak open automaton derived from $A$.
%% Suppose we have the following weak open transitions in $\WT$:
%% \[\openrule
%%          {
%%            \overline{\gamma_1},\Pred_1,\Post_1   }
%%          {s \OTWeakarrow {\tau} s_1}
%% \qquad 
%% \openrule
%%          {
%%            \overline{\gamma_2},\Pred_2,\Post_2   }
%%          {s_1 \OTWeakarrow {\alpha} s_2}
%% \qquad
%% \openrule
%%          {
%%            \overline{\gamma_3},\Pred_3,\Post_3   }
%%          {s_2 \OTWeakarrow {\tau} s_3}
%% \]
%% Then there is a weak  open transition of the following form in $\WT$:
%% \[\openrule
%%          {
%%            \overline{\gamma},\Pred,\Post   }
%%          {s \OTWeakarrow {\alpha} s_3}
%% \]
%% Suppose additionally $\Post_3=({x_k\gets e_k})^{k\in K}$, we have the following
%% \[
%% \begin{array}{rcl}
%% \gamma&=& \gamma_1\dotcup\gamma_2\dotcup\gamma_3\\
%% \Pred&=&\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}\\
%% \Post &=& ({x_k\gets ((e_k\subst{\Post_2})\subst{\Post_1})})^{k\in K}
%% \end{array}\]
%% \end{property}

@@

\subsection{Weak bisimulation for open pNets}
\begin{definition}[Non-observability of silent actions for pNets]~\\
A pNet $\mylangle \pNet_i^{i\in I} , \Sort_j^{j\in J}, \overline{\symb{SV}}\myrangle$
 \emph{cannot observe silent actions} if it verifies:\\ $\forall i\in I\uplus J.\, \SV{(i\mapsto \tau)}{\tau}{\symb{True}}\in \overline{\symb{SV}}$ and $\forall SV_k\in \overline{\symb{SV}}, \forall i,j\in I_k\uplus J_k.\,i\neq j \Rightarrow \alpha'_k\neq\tau$
\end{definition}

@@

\begin{definition}[Notation: Semantics of pNets as a weak open automaton]
Let $A$ be the open automaton expressing the semantics of an open \pNet: $\pNet$; let $<J,\mathcal{S},s_0,  \WT>$ be the weak open automata derived from $A$. For any $\WOT\in\WT$, we denote $\pNet \models \WOT$, this defines the semantics of the pNet $\pNet$ as a weak open automaton.
\end{definition}

%% \TODO{move text below}

%% \begin{property}[Decomposition of weak open transitions] \TODO{TODO}
%% Let $A$ be the open automaton that is the semantics of a \pNet $P$, and $<J,\mathcal{S},s_0,\WT>$ a weak open automaton derived from $A$.
%% Suppose we have the following weak open transitions in $\WT$:
%% \[\openrule
%%          {
%%            \overline{\gamma_1},\Pred_1,\Post_1   }
%%          {s \OTWeakarrow {\tau} s_1}
%% \qquad 
%% \openrule
%%          {
%%            \overline{\gamma_2},\Pred_2,\Post_2   }
%%          {s_1 \OTWeakarrow {\alpha} s_2}
%% \qquad
%% \openrule
%%          {
%%            \overline{\gamma_3},\Pred_3,\Post_3   }
%%          {s_2 \OTWeakarrow {\tau} s_3}
%% \]
%% Then there is ag weak  open transition of the following form in $\WT$:
%% \[\openrule
%%          {
%%            \overline{\gamma},\Pred,\Post   }
%%          {s \OTWeakarrow {\alpha} s_3}
%% \]
%% Suppose additionally $\Post_3=({x_k\gets e_k})^{k\in K}$, we have the following
%% \[
%% \begin{array}{rcl}
%% \gamma&=& \gamma_1\dotcup\gamma_2\dotcup\gamma_3\\
%% \Pred&=&\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}\\
%% \Post &=& ({x_k\gets ((e_k\subst{\Post_2})\subst{\Post_1})})^{k\in K}
%% \end{array}\]
%% \end{property}

\TODO{We will discuss this corollary and the title.}
\begin{corollary}[Weak Composability]
	Consider two weak FH-bisimilar pNets with an arbitrary number of holes, when replacing, 
	inside those two original pNets, a subset of the holes by weak FH-bisimilar pNets, we 
	obtain two weak FH-bisimilar pNets.
\end{corollary}

@@

\TODO{Here we should include:
  %% \\  - A sufficient condition for an open PNet to preserve Weak-FH,
  %% including: 1) no non-deterministic Tau-choice in the pLTSs, 2) a
  %% synch vector propagating Tau from a subnet $pNet_i$ mandatory
  %% everytime $pNet_i$ can do some visible action in the current
  %% context, 3) if we have a vector <Tau, -, a, - > --> b with ``a'' a visible
  %% action of some CLOSED subnet, then b=Tau. 
  \\  - the corresponding Congruence theorem
  \\- An example
}
